---
title: "simpleRegress"
author: "Kaylie Fagoaga"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

For model length ~ neck, adjusted $r^2$ = 0.75

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## Length based on Age 
```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```

For model length ~ age, adjusted $r^2$ = 0.51, this is less than the adjusted $r^2$ for neck.       
This model is not as good as the neck model

## Length ~ Chest  
```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```

For model length ~ chest, adjusted $r^2$ = 0.79, this is greater than the adjusted $r^2$ of 0.75 for neck.    
This model is better than the neck model. 

## Length ~ Sex
```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```

For model length ~ sex, adjusted $r^2$ = -0.008373, this is less than the adjusted $r^2$ of 0.79 for chest   
This is not as good as the chest model.  

## Length ~ Headlength
```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)
```

For model length ~ headlength, adjusted $r^2$ = 0.84, this is greater than the adjusted $r^2$ of 0.79 for chest.       
This model is better than chest model.   

## Length ~ Headwidth
```{r}
headwthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwthModel)
summary.lm(headwthModel)
```

For model length ~ headwith, adjusted $r^2$ = 0.53, this is less than adjusted $r^2$ of 0.84 for headlength.        
This is not as good as the headlength model.   

## Length ~ Weight
```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```

For model length ~ weight, adjusted $r^2$ = 0.74, this is less than adjusted $r^2$ of 0.84 for headlength.        
This is not as good as the headlength model. 







